Tropical Mathematics for the Lambda Calculus

### 2023-01-17 section3C.tex
commit

%A model of the $\STDLC$ is usually understood as so-called \emph{Cartesian closed differential categories} (CC$\partial$C), see \cite{Manzo2010} for details.
%In order to treat the $+$ and the constructor $D[\_]\cdot (\_)$ of $\STDLC$, the main features of a CC$\partial$C $\C C$ are that:
%
%1) $\C C$ is a left-additive-CCC, i.e.\ its Homsets are commutative monoids and its Cartesian closed structure is well behaved w.r.t.\ this monoid structure;
%
%2) $\C C$ is equipped with a differential operator map $D:\HOM{\C C}{X}{Y}\to \HOM{\C C}{X\times X}{Y}$ (here $\times$ is the Cartesian product of $\C C$) satisfying $8$ axioms, called D1, ..., D7, D-curry.

@@

There is more: not only $\LREL_!$ is a model of the $\STDLC$, but this model is also well-behaved w.r.t.\ to the Taylor expansion of $\lam$-terms.
This is expressed by the following property (see [Definition 4.22, \cite{Manzo2010}], where it is called the fact of ``modeling the Taylor expansion'').

\begin{theorem}\label{thm:modelsTaylor}
 In $\LREL_!$ equipped with $D$, all morphisms can be Taylor expanded, i.e.\ in $\LREL_!$ we have that, for all
$t\in\HOM{\LREL_!}{Z}{X\multimap Y}$ and $s\in\HOM{\LREL_!}{Z}{X}$, the evaluation of $t$ over $s$ yields 
 \begin{align}\label{eq:taylorcat}
  \RM{ev}\circ_!\langle t,s\rangle =
  \inf\limits_{n\in\N}
  \set{((\dots((\Lambda^- t)\star s)\star \dots)\star s)\circ_! \langle \RM{id},\infty \rangle}.
 \end{align}
 \end{theorem}
It is worth discussing the formula above a bit more. 
Here, $\Lambda^-$
%:\HOM{\LREL}{!Z}{X\multimap Y}\to \HOM{\LREL}{!(Z+X)}{Y}$
 is the uncurry operator and 
%$\star:\HOM{\LREL}{!(Z+X)}{Y}\times\HOM{\LREL}{!Z}{X}\to \HOM{\LREL}{!(Z+X)}{Y}$ is defined as 
$u\star s= (Du)\circ \langle s,\infty\rangle$ corresponds to linear application of $s$ to $u$. Hence the right-hand term in \eqref{eq:taylorcat} corresponds to the $\inf$ of the $n$-th derivative of $\Lambda^{-}t$ applied to ``$n$ copies'' of $s$,  i.e.~it coincides with the tropical interpretation of the Taylor formula \eqref{eq:taylor}.
In other words, \eqref{eq:taylorcat} says that the series corresponding to the Taylor expansion of an ordinary $\lam$-term $M$ converges to $M$ [Theorem 4.23, \cite{Manzo2010}].

>>>>

%A model of the $\STDLC$ is usually understood as so-called \emph{Cartesian closed differential categories} (CC$\partial$C), see \cite{Manzo2012} for details.
%In order to treat the $+$ and the constructor $D[\_]\cdot (\_)$ of $\STDLC$, the main features of a CC$\partial$C $\C C$ are that:
%
%1) $\C C$ is a left-additive-CCC, i.e.\ its Homsets are commutative monoids and its Cartesian closed structure is well behaved w.r.t.\ this monoid structure;
%
%2) $\C C$ is equipped with a differential operator map $D:\HOM{\C C}{X}{Y}\to \HOM{\C C}{X\times X}{Y}$ (here $\times$ is the Cartesian product of $\C C$) satisfying $8$ axioms, called D1, ..., D7, D-curry.

@@

There is more: not only $\LREL_!$ is a model of the $\STDLC$, but this model is also well-behaved w.r.t.\ to the Taylor expansion of $\lam$-terms.
This is expressed by the following property (see [Definition 4.22, \cite{Manzo2012}], where it is called the fact of ``modeling the Taylor expansion'').

\begin{theorem}\label{thm:modelsTaylor}
 In $\LREL_!$ equipped with $D$, all morphisms can be Taylor expanded, i.e.\ in $\LREL_!$ we have that, for all
$t\in\HOM{\LREL_!}{Z}{X\multimap Y}$ and $s\in\HOM{\LREL_!}{Z}{X}$, the evaluation of $t$ over $s$ yields 
 \begin{align}\label{eq:taylorcat}
  \RM{ev}\circ_!\langle t,s\rangle =
  \inf\limits_{n\in\N}
  \set{((\dots((\Lambda^- t)\star s)\star \dots)\star s)\circ_! \langle \RM{id},\infty \rangle}.
 \end{align}
 \end{theorem}
It is worth discussing the formula above a bit more. 
Here, $\Lambda^-$
%:\HOM{\LREL}{!Z}{X\multimap Y}\to \HOM{\LREL}{!(Z+X)}{Y}$
 is the uncurry operator and 
%$\star:\HOM{\LREL}{!(Z+X)}{Y}\times\HOM{\LREL}{!Z}{X}\to \HOM{\LREL}{!(Z+X)}{Y}$ is defined as 
$u\star s= (Du)\circ_{!} \langle \langle  \infty, s\circ_{!} \pi_{1}\rangle,\mathrm{id}\rangle$ corresponds to linear application of $s$ to $u$. Hence the right-hand term in \eqref{eq:taylorcat} corresponds to the $\inf$ of the $n$-th derivative of $\Lambda^{-}t$ applied to ``$n$ copies'' of $s$,  i.e.~it coincides with the tropical interpretation of the Taylor formula \eqref{eq:taylor}.
In other words, \eqref{eq:taylorcat} says that the series corresponding to the Taylor expansion of an ordinary $\lam$-term $M$ converges to $M$ [Theorem 4.23, \cite{Manzo2012}].

### 2023-01-17 section3intro.tex
commit

 As it is expected, $Q^X$ is a $Q$-semimodule and the bijection $\hat{(\cdot)}$ identifies $\HOM{\QREL}{X}{Y}$ with the set of linear maps from $Q^X$ to $Q^Y$ via
 \begin{equation}
 \hat t(x)_a:=\sum\limits_{b\in Y} t_{a,b}x_a
 \end{equation}

 \begin{remark}
 Following \cite{Manzo2013, Hofmann2014, Ehrhard2005}, we 
 chose to see a matrix from $X$ to $Y$ as a map $s:X\times Y\to Q$.
% 
% fix $\HOM{\QREL}{X}{Y}:=Q^{X\times Y}$ with composition $st:X\times Z\to Q$ of $s:Y\times Z\to Q$ and $t:X\times Y\to Q$ defined by $(st)_{a,c}:=\sum\limits_{b\in Y} s_{b,c}t_{a,b}$.
In linear algebra, a map $X\times Y\to Q$ is usually called a ``$Y\times X$-matrix'', meaning $Y$ rows and $X$ columns.
In particular, the product of such a matrix for a vector defines a map $Q^Y\to Q^X$, hence the opposite of what shown above.
\end{remark}
%
%In $\QREL^{op}$ (which corresponds to systematically taking transpose matrices), composition coincides with the product matrix/matrix and $\hat{(\cdot)}$ with the product matrix/vector.
%In order to avoid confusion, we will refer to a $t\in Q^{X\times Y}$ just as a \emph{matrix from $X$ to $Y$}.
%
%
%
%We must fix a convention for matrices: following \cite{Manzo2013, Hofmann2014, Ehrhard2005}, we fix $\HOM{\QREL}{X}{Y}:=Q^{X\times Y}$ with composition $st:X\times Z\to Q$ of $s:Y\times Z\to Q$ and $t:X\times Y\to Q$ defined by $(st)_{a,c}:=\sum\limits_{b\in Y} s_{b,c}t_{a,b}$.
%In linear algebra, a map $X\times Y\to Q$ is usually called a ``$Y\times X$-matrix'', meaning $Y$ rows and $X$ columns.
%In particular, the product of such a matrix for a vector defines a map $Q^Y\to Q^X$.
%Instead, we prefer to see a $t\in\HOM{\QREL}{X}{Y}$ as giving rise to a map $\hat t:Q^X\to Q^Y$ defined by $\hat t(x)_a:=\sum\limits_{b\in Y} t_{a,b}x_a$.
%In $\QREL^{op}$ (which corresponds to systematically taking transpose matrices), composition coincides with the product matrix/matrix and $\hat{(\cdot)}$ with the product matrix/vector.
%In order to avoid confusion, we will refer to a $t\in Q^{X\times Y}$ just as a \emph{matrix from $X$ to $Y$}.
%As it is expected, $Q^X$ is a $Q$-semimodule and the bijection $\hat{(\cdot)}$ identifies $\HOM{\QREL}{X}{Y}$ with the set of linear maps from $Q^X$ to $Q^Y$.

>>>>

 As it is expected, $Q^X$ is a $Q$-semimodule and the bijection $\hat{(\cdot)}$ identifies $\HOM{\QREL}{X}{Y}$ with the set of linear maps from $Q^X$ to $Q^Y$ via
 \begin{equation}
 \hat t(x)_b:=\sum\limits_{a\in X} t_{a,b}x_a
 \end{equation}

 \begin{remark}
 Following \cite{Manzo2013, Hofmann2014, Ehrhard2005}, we 
 chose to see a matrix from $X$ to $Y$ as a map $s:X\times Y\to Q$.
% 
% fix $\HOM{\QREL}{X}{Y}:=Q^{X\times Y}$ with composition $st:X\times Z\to Q$ of $s:Y\times Z\to Q$ and $t:X\times Y\to Q$ defined by $(st)_{a,c}:=\sum\limits_{b\in Y} s_{b,c}t_{a,b}$.
Notice that usual linear algebra conventions correspond to working in $\LREL^{\op}$: a matrix $X\times Y\to Q$ is usually called a ``$Y\times X$-matrix'', meaning $Y$ rows and $X$ columns, and the usual matrix-vector product defines a map $Q^Y\to Q^X$.
\end{remark}
%
%In $\QREL^{op}$ (which corresponds to systematically taking transpose matrices), composition coincides with the product matrix/matrix and $\hat{(\cdot)}$ with the product matrix/vector.
%In order to avoid confusion, we will refer to a $t\in Q^{X\times Y}$ just as a \emph{matrix from $X$ to $Y$}.
%
%
%
%We must fix a convention for matrices: following \cite{Manzo2013, Hofmann2014, Ehrhard2005}, we fix $\HOM{\QREL}{X}{Y}:=Q^{X\times Y}$ with composition $st:X\times Z\to Q$ of $s:Y\times Z\to Q$ and $t:X\times Y\to Q$ defined by $(st)_{a,c}:=\sum\limits_{b\in Y} s_{b,c}t_{a,b}$.
%In linear algebra, a map $X\times Y\to Q$ is usually called a ``$Y\times X$-matrix'', meaning $Y$ rows and $X$ columns.
%In particular, the product of such a matrix for a vector defines a map $Q^Y\to Q^X$.
%Instead, we prefer to see a $t\in\HOM{\QREL}{X}{Y}$ as giving rise to a map $\hat t:Q^X\to Q^Y$ defined by $\hat t(x)_a:=\sum\limits_{b\in Y} t_{a,b}x_a$.
%In $\QREL^{op}$ (which corresponds to systematically taking transpose matrices), composition coincides with the product matrix/matrix and $\hat{(\cdot)}$ with the product matrix/vector.
%In order to avoid confusion, we will refer to a $t\in Q^{X\times Y}$ just as a \emph{matrix from $X$ to $Y$}.
%As it is expected, $Q^X$ is a $Q$-semimodule and the bijection $\hat{(\cdot)}$ identifies $\HOM{\QREL}{X}{Y}$ with the set of linear maps from $Q^X$ to $Q^Y$.

### 2023-01-17 section4.tex
commit

A tropical polynomial, as we saw, is a piece-wise linear function of the form $\varphi(x)=\min_{i=1,\dots,k}\{ix+c_{i}\}$. For instance, the polynomials $\varphi_{n}(x)=\min_{i\leq n}\{in+2^{-i}\}$
are plotted in Fig.~\ref{fig:plot1} for $n=2,3,4$.
A \emph{tropical root} of a tropical polynomials $\varphi$ is a point $x\in \Lawv$ where $\varphi$ is not differentiable. In other words, the roots of $\varphi$ correspond with the points where the slope of $\varphi$ changes (i.e.~the points where the minimum defining $\varphi$ is attained twice). A simple calculation shows that this condition corresponds, in the tropical setting, to the usual notion of root.

@@

\end{axis}
\end{tikzpicture}
\caption{}
\label{fig:plot2}
\end{subfigure}
\label{fig:plot12}
\caption{
%\ref{fig:plot1}
 (a) Plot of the tropical polynomials $\varphi_{2}$ (in red), $\varphi_{3}$ (in blue), $\varphi_{4}$ (in orange) and of their limit tLs $\varphi$ (in violet).\\
%\ref{fig:plot2}
(b) Plot of
$\varphi(x,y)=\min\{2x, 2x+y,3y\}$.}
\end{figure}

%
%
%We will take as our running example
%%given respectively by Equation~\ref{eq:polytrop} and Equation~\ref{eq:defTLS}.
%A positive $x\in[0,\infty)$ is said to be a (finite) \emph{tropical root} of a tropical Laurent series $f:\Lawv\to\Lawv$ iff $f$ is not differentiable at $x$ (w.r.t.\ the usual topology on $\BB R_{\geq0}$.
%This is equivalent to ask that the $\inf$ defining $f$ at $x$ is obtained \emph{at least} twice.{\color{red} \`e vero anche per TLS o solo per poly questo? Inoltre, Robol parla anche di finite end-points of the domain: nel nostro caso sarebbe $x=0$}

@@

%In Example~\ref{ex:famous_ex}, 
Again, by looking at Fig~\ref{fig:plot1}, it appears that $\varphi$ behaves \emph{locally} like the polynomials $\varphi_{n}$. In particular, for all $\epsilon >0$, $f$ coincides on $[\epsilon,\infty]$ with some polynomial $\varphi_{n}$. Indeed, we can compute  by hand that $\varphi$ coincides with $\varphi_{n}$ 
for $\epsilon \geq \log2/2^{n+1}$.
However, at
%
%$x\in [\epsilon,\infty]$, with $\epsilon>0$
%It can be proven by hand that $\varphi(x)$ is a $\min$ for all $x>0$.
 $x=0$ we have that $\varphi(x=0)=\inf\limits_{n\in\N} \frac{1}{2^n}=0$, and this is the only point where the $\inf$ is \emph{not} a $\min$.
Also, while the derivative of $f$ is bounded on all $\BB R_{>0}$, at $x=0$ it tends to $\infty$.
This phenomenon is reminiscent of [Example 7, \cite{Ehrhard2005}],
%Differentials and Distances in Probabilistic Coherence Spaces. FSCD 2019
which actually motivated our first investigations.
In fact, these properties are shared by all tLs with \emph{finitely} many variables, as shown by the following result.

Remark that $!\set{1,\dots,k}$ can be identified with $\N^k$.
So the matrix of a tLs $f$ with finitely many variables $x=x_1,\dots,x_k$ (and one variable $f(x)$ in output) can be given as a $\widehat f:\N^k\to\Lawv$, and $f$ has shape $f(x)=\inf\limits_{n\in \N^k}\set{nx+\widehat f(n)}$, where $nx$ is the scalar product.

@@

\subsection{$\Lawv^{X}$ as a normed cone}

@@

\subsection{Metric results}%$\Lawv^{X}$ as a metric space.}

>>>>

A tropical polynomial, as we saw, is a piece-wise linear function of the form $f(x)=\min_{i_{1},\dots, i_{k}}\{i_{j}x+c_{i_{j}}\}$. For instance, the polynomials $\varphi_{n}(x)=\min_{i\leq n}\{in+2^{-i}\}$
are plotted in Fig.~\ref{fig:plot1} for $n=2,3,4$.
A \emph{tropical root} of a tropical polynomials $\varphi$ is a point $x\in \Lawv$ where $\varphi$ is not differentiable. In other words, the roots of $\varphi$ correspond with the points where the slope of $\varphi$ changes (i.e.~the points where the minimum defining $\varphi$ is attained at least twice). 
It can be shown that this corresponds to the usual factorization property of roots: for any root $x_{0}$, $f$ factorizes as
$f(x)=\min\{x,x_{0}\}+ g(x)$.
%
%A
% simple calculation shows that this condition corresponds, in the tropical setting, to the usual notion of root.

@@

\end{axis}
\end{tikzpicture}
\caption{}
\label{fig:plot2}
\end{subfigure}
\label{fig:plot12}
\caption{
%\ref{fig:plot1}
 (a) Plot of the tropical polynomials $\varphi_{2}$ (in red), $\varphi_{3}$ (in blue), $\varphi_{4}$ (in orange) and of their limit tLs $\varphi$ (in violet). The points where the slope changes are  the tropical roots of $\varphi$, i.e.~the points $x$ where $ix+2^{-i}=(i+1)x+2^{-i+1}$, given by $x=2^{-i+1}$.
%\ref{fig:plot2}
(b) Plot of
$\varphi(x,y)=\min\{2x, 2x+y,3y\}$.}
\end{figure}

%
%
%We will take as our running example
%%given respectively by Equation~\ref{eq:polytrop} and Equation~\ref{eq:defTLS}.
%A positive $x\in[0,\infty)$ is said to be a (finite) \emph{tropical root} of a tropical Laurent series $f:\Lawv\to\Lawv$ iff $f$ is not differentiable at $x$ (w.r.t.\ the usual topology on $\BB R_{\geq0}$.This is equivalent to ask that the $\inf$ defining $f$ at $x$ is obtained \emph{at least} twice.{\color{red} \`e vero anche per TLS o solo per poly questo? Inoltre, Robol parla anche di finite end-points of the domain: nel nostro caso sarebbe $x=0$}

@@

%In Example~\ref{ex:famous_ex}, 
Again, by looking at Fig~\ref{fig:plot1}, it appears that $\varphi$ behaves \emph{locally} like the polynomials $\varphi_{n}$. In particular, for all $\epsilon >0$, $f$ coincides on $[\epsilon,\infty]$ with some polynomial $\varphi_{n}$. Indeed, we can compute  by hand that $\varphi$ coincides with $\varphi_{n}$ 
for $\epsilon \geq \log2/2^{n+1}$.
However, at
%
%$x\in [\epsilon,\infty]$, with $\epsilon>0$
%It can be proven by hand that $\varphi(x)$ is a $\min$ for all $x>0$.
 $x=0$ we have that $\varphi(x=0)=\inf_{n\in\N} \frac{1}{2^n}=0$, and this is the only point where the $\inf$ is \emph{not} a $\min$.
Also, while the derivative of $f$ is bounded on all $\BB R_{>0}$, at $x=0$ it tends to $\infty$.
This phenomenon is reminiscent of [Example 7, \cite{Ehrhard2005}],
%Differentials and Distances in Probabilistic Coherence Spaces. FSCD 2019
which actually motivated our first investigations.
In fact, these properties are shared by all tLs with \emph{finitely} many variables, as shown by the following result.

Remark that $!\set{1,\dots,k}$ can be identified with $\N^k$.
So the matrix of a tLs $f$ with finitely many variables $x=x_1,\dots,x_k$ (and one variable $f(x)$ in output) can be given as a $\widehat f:\N^k\to\Lawv$, and $f$ has shape $f(x)=\inf_{n\in \N^k}\set{nx+\widehat f(n)}$, where $nx$ is the scalar product.

@@

\subsection{Continuity of tLs}%$\Lawv^{X}$ as a normed cone}

@@

\subsection{Lipschitz-Continuity of tLs}%$\Lawv^{X}$ as a metric space.}

### 2023-01-17 section5.tex
commit

Secondly, the Lipschitz conditions studied in the previous section allow us to reason on program distances in a \emph{compositional} way: suppose, as before, that $M,N:A$ are two programs such that $M$ improves on $N$ by $\epsilon$, and let $\TT C[-]:A \to \B{nat}$ indicate a context; knowing that the interpretation of $\TT C$ is $k$-Lipschitz-continuous on some open set containing both $\model M$ and $\model N$, allows us to immediately deduce that $\TT C[M]$ improves on $\TT C[N]$ by $k \epsilon$. 
Observe that this will typically be the case when the Taylor expansions of $\TT C[M]$ and $\TT C[N]$ actually yields a \emph{finite} sum of at most $k$ terms, i.e.~when 
\begin{align}
\TT C[M] = \sum_{i=0}^{k} \TT D^{(k)}\Big[\lambda x.\TT C[x]\Big]\cdot M^{k}
\end{align}
and similarly for $\TT C[N]$. It is here worth recalling that, for $\STDLC$, a well-known result \cite{difflambda} is that the Taylor expansion of a closed application $MN$ is always finite, although its non-zero coefficients may be arbitrarily high. 
Notably, these observations suggest to study tropical versions of \emph{finiteness spaces} \cite{Ehrhard2005}, 
a variant of the relational semantics modeling strongly normalizing programs via \emph{finite} power series.
We mention this point in the Section~\ref{sec:concl}

@@

\subsection{Resource Analysis for Differential Privacy}

The typical situation in differential privacy is where one considers a probabilistic query $f: \mathsf{db}\to [0,1]^{X}$ over some database, and one requires that $f$ should not be \emph{too sensitive} to small changes in the output, in other to prevent potential leaks of private information about individual items in $\mathsf{db}$ (for instance, an element $x\in\mathsf{db}$ could be the list of students of some university and $f(x)$ indicates the percentage of female students$\mathsf{db}$).

More formally, differential privacy is defined as follows \cite{Reed2010}:
\begin{definition}
Let $f: \mathsf{db}\to [0,1]^{X}$ and $\epsilon \in \BB R_{\geq 0}$. $f$ is said \emph{$\epsilon$-differentially private} if for all $x,x'\in \mathsf{db}$
differing by $L$ items, for all $s\in X$, 
$$
f(x)(s) \ = \ e^{\epsilon L} f(x')(s)
$$
\end{definition}

A well-studied approach to ensure differential privacy for higher-order programs is to use bounded type systems like $\mathsf{Fuzz}$ \cite{Reed2010}. Indeed, such systems come with an \emph{a priori} warrant that well-typed programs are Lipschitz.

The use of tropical semantics suggests how resource analysis could also be used to provide bounds for differential privacy. 
Suppose our probabilistic query $f$ can be expressed as a power series (this is what happens e.g.~in \emph{probabilistic coherent spaces} \cite{Ehrhard2011}). Then, if we discover, either by studying differential properties of $f$, or using methods from convex analysis as suggested in the previous section (e.g.~Theorem \ref{??}), 
 that the tropicalization $\trop f$ satisfies a Lipschitz condition, we may use this fact to deduce that $f$ is differentially private, as shown by the result below.

%
%We have seen how a higher-order probabilistic program, which is expressed as a polynomial, can be tropicalized. This operation can be extended to programs expressed as power series by taking the limit. 
%
\begin{proposition}
Let $f: [0,1]^{X}\to [0,1]^{Y}$ be an analytic function. If $\trop f$ is Lipschitz over some open set $U$, then $f$ is differentially private over $\psi_{t}(U)$, for small enough $t$.
\end{proposition}
\begin{proof}
We consider the case $X=Y=\B 1$ for simplicity. 
Express $f$ as a power series $f(x)=\sum_{n}\psi_{e}(\widehat f_{n})x^{n}$, so that $\trop f(\alpha)= \inf_{n}\left \{
n\alpha + \widehat f_{n}
\right\}$.
Let us define a family of intermediate functions $\trop_{t}f(\alpha)=\phi_{t}(\stackrel{t}{\sum_{n}}\widehat{f}_{n}\times_{t}n\alpha)$. These, by construction, satisfy
\begin{align}\label{eq:phit}
f(\phi_{t}(\alpha)) = \phi_{t}(\trop_{t}f(\alpha))
\end{align}
and, using the fact that $\alpha\sumt{t}\beta\to\min\{\alpha,\beta\}$ and $\alpha\prodt{t}\beta\to \alpha+\beta$, we can deduce that $\trop f(\alpha)=\lim_{t\to 0}\trop_{t}f(\alpha)$ (for this, we use the fact that for all $m\in \BB N$,
$\min_{i\leq m}\alpha_{i}\leq \stackrel{t}{\sum}_{i\leq m}\alpha_{i}\leq 
(\min_{i\leq m}\alpha_{i})+t\log m$).

Now, using the fact that $\trop f$ and $\trop_{t} f$ come arbitrarily close with $t$ small enough, together with \eqref{eq:phit}, the Lipschitz condition  
$|\trop f(\phi_{t}(x))-\trop f(\phi_{t}(y))|\leq L\cdot |\phi_{t}(x)-\phi_{t}(y)|$ translates into the differential privacy condition 
$f(x) \leq e^{L\cdot |\log x-\log y|} f(y)$ (supposing $y\geq x$ and $f(y)\neq 0$).
\end{proof}

%General discussion: optimization properties behave in a Lipschitz way.
%
%
%- differential privacy and Lipschitzness
%
%
%- log-probabilities and tropical roots 
%
%
%- counting computation steps (from Manzonetto, but add relational ``Lipschitz'' reasoning)
%
%
%- measuring duplications of discrete functions (needs finiteness!)

>>>>

Secondly, the Lipschitz conditions from Section \ref{section4} allow us to reason on program distances in a \emph{compositional} way: suppose, as before, that $M,N:A$ are two programs such that $M$ improves on $N$ by $\epsilon$, and let $\TT C[-]:A \to \B{nat}$ indicate a context; knowing that the interpretation of $\TT C$ is $k$-Lipschitz-continuous on some open set containing both $\model M$ and $\model N$, allows us to immediately deduce that $\TT C[M]$ improves on $\TT C[N]$ by $k \epsilon$. 
Observe that this will typically be the case when the Taylor expansions of $\TT C[M]$ and $\TT C[N]$ actually yields a \emph{finite} sum of at most $k$ terms, i.e.~when 
\begin{align}
\TT C[M] = \sum_{i=0}^{k} \TT D^{(k)}\Big[\lambda x.\TT C[x]\Big]\cdot M^{k}
\end{align}
and similarly for $\TT C[N]$. It is here worth recalling that, for $\STDLC$, a well-known result \cite{difflambda} is that the Taylor expansion of a closed application $MN$ is always finite, although its non-zero coefficients may be arbitrarily high. 
Notably, these observations suggest to study tropical versions of \emph{finiteness spaces} \cite{Ehrhard2005}, 
a variant of the relational semantics modeling strongly normalizing programs via \emph{finite} power series.
We mention this point in Section~\ref{section8}

@@
